---
title: "Cropping data recorded before deployment and after retrieval"
output:
  html_document:
    theme: flatly
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 3
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Editable script

<!-- this script crops raw data from HOBO files -->

<!-- input: raw data files (csv format), ldrtimes file (Excel format) -->
<!-- output: cropped data files (csv format), plots of cropped data -->

<!-- Important: -->
<!--   Before running this script, rename raw HOBO files to have filenames in -->
<!--     this format: sitename_medium_deployseason_deployyear.csv -->
<!--     e.g. NolanLower_air_sum_23.csv -->

<!--   Make sure ldrtimes has columns for deployment season and year: -->
<!--   ldrtimes$deploy_season = "sum" -->
<!--   ldrtimes$deploy_year = "23" -->

<!-- we can label the red lines on the plots if you like -->

<!-- before running this script, update the file paths below -->
<!-- - to toggle a line between commented (#...) and uncommented (...), use -->
<!--   Shift-Control-C (Windows) or Shift-Command-C (Mac) -->
<!-- - make sure to add a slash "/" at the end of each file path! -->

<!-- if you'd like to have your computer make sound when the job is done, -->
<!-- install and load this package -->
<!-- install.packages("beepr") # uncomment and run this step only once, then comment it again -->
<!-- library(beepr) # leave this uncommented -->
<!-- beep(8) # uncomment if you want to use beepr! -->
<!-- system("say 'Attention all your data has been cropped'") # or on a Mac you can use this -->

<!-- base path: this is the shared part of the filepath to the directories -->
<!-- containing your raw data files, cropped data files, and cropped data plots -->
<!-- - if these folders are in the dataQCtools R project folder, you can use a relative path -->
<!-- - otherwise you can make this an absolute path -->
<!-- - you can also delete base_loc and paste0(base_loc, ...) and just explicitly -->
<!--   specify rawdata_loc, cropped_loc, and croppedplots_loc (see make_qc_plots.R -->
<!--   for an example of this) -->
<!-- - make sure filepaths for folders always end in "/"! -->

Here is a script you can use if you would prefer to have all the code in a single script rather than an installable package format. Just click the clipboard icon in the top right corner of this code chunk to copy the code, then open a new R script in RStudio and paste it there, and save the script wherever you would like to store it on your computer.

```{r, eval = FALSE}
# load necessary libraries
library(zoo)
library(grid)
library(plotly)
library(readxl)
library(scales)
library(runner) # for moving averages
library(ggthemes) # for plotting
library(tidyverse) # need for dplyr and ggplot2
library(lubridate) # for handling dates
library(gridExtra)
library(flextable) # for making tables with conditional formatting, merged cells and more
library(RColorBrewer)

base_loc <- "data/2022_summer/"

# directory containing your raw data
rawdata_loc <- paste0(base_loc, "1_raw_csv/")

# directory where your cropped data will be (or currently is) stored
cropped_loc <- paste0(base_loc, "2_cropped_csv/")

# directory where plots of cropped and raw data will be stored
croppedplots_loc <- paste0(base_loc, "2_cropped_plots/")

# name of your LDRTimes file that has the lookup table with deployment/retrieval dates
ldrtimes_fn <- "LDRTimes_summer22.xlsx"

# check that R can find your raw data files
# Get all temperature data filenames
# note: * is called a glob, short for global
# IMPORTANT: filenames must be in the format sitename_medium_deployseason_deployyear.csv
#            e.g. NolanLower_air_sum_23.csv
csv_files = list.files(path = rawdata_loc, pattern = '*csv')
# this is now a list of all filenames; we haven't read in the data yet, but
# make sure this lists all the raw files you want to crop
csv_files

# read in LDR file and take a look at it
# note: this assumes the LDR file is in the folder indicated by rawdata_loc
ldrtimes = read_xlsx(paste0(rawdata_loc, ldrtimes_fn))

# once you're sure that file paths are working and your ldrtimes looks right,
# crop the files!

# get a list of all the temperature data filenames
csv_files = list.files(path = rawdata_loc, pattern = '*csv')
# read in LDR file
# - this assumes the LDR file is an excel file in the folder indicated by rawdata_loc
ldrtimes = read_xlsx(paste0(rawdata_loc, ldrtimes_fn))

i = 0
for(this.file in csv_files){
  i = i + 1
  #this.file = csv_files[1] # uncomment to troubleshoot within loop
  cat(paste0("Reading file ", i, " of ", length(csv_files), ": ", this.file), fill = TRUE)
  
  # extract metadata from the filename
  filename.parts = strsplit(this.file, '[_.]')[[1]]
  csv.site = filename.parts[1]
  csv.media = filename.parts[2]
  csv.season = filename.parts[3]
  csv.year = filename.parts[4]
  
  # convert the character-format datetime to an R POSIXct object
  # ymd_hm is the format the character string is in initially; it tells R
  # how to read and interpret the character string
  # sometimes R reads in the datetime format as mdy_hms and sometimes mdy_hm. 
  # This tryCatch handles either hh:mm:ss or hh:mm format in csv files
  this.data =  tryCatch(
    { 
      read_csv(paste0(rawdata_loc, this.file),
               skip = 2, # skip the first two lines of the file
               col_select = 1:3, # read only the first three columns of data
               col_names = FALSE, # don't try to name columns from a row of the file
               show_col_types = FALSE) %>% # suppresses print message
        rename("row.num" = X1,
               "datetime" = X2,
               "temperature" = X3) %>%
        mutate(datetime = mdy_hms(datetime)) #for datetime in hh:mm:ss
    }, 
    warning = function(cond) { #if datetime isn't in hh:mm:ss, will now try hh:mm format
      read_csv(paste0(rawdata_loc, this.file),
               skip = 2, # skip the first two lines of the file
               col_select = 1:3, # read only the first three columns of data
               col_names = FALSE, # don't try to name columns from a row of the file
               show_col_types = FALSE) %>% # suppresses print message
        rename("row.num" = X1,
               "datetime" = X2,
               "temperature" = X3) %>%
        mutate(datetime = mdy_hm(datetime)) #for datetime in hh:mm
    }
  )
  
  # crop the data
  deploy.retrieval = ldrtimes %>%
    # select the row(s) of ldrtimes that match this datafile
    # should be exactly one row, but if there are no rows or multiple rows that
    # match, this step will pull that many rows
    filter(site == csv.site, deploy_season == csv.season,
           deploy_year == csv.year, media == csv.media) %>%
    # keep just the deploy_time and retrieval_time variables/columns
    select(deploy_time, retrieval_time)
  
  if(nrow(deploy.retrieval) == 0){
    stop("no rows of ldrtimes matched this csv file.")
  }
  if(nrow(deploy.retrieval) > 1){
    stop("multiple rows of ldrtimes matched this csv file.")
  }
  
  deploy = deploy.retrieval$deploy_time
  retrieval = deploy.retrieval$retrieval_time
  
  if(retrieval > deploy) {
    cropped.data = filter(this.data,
                          datetime > deploy,
                          datetime < retrieval)
    
  } # if(retrieval > deploy)
  
  # write cropped csv files to cropped folder
  write_csv(cropped.data,
            file=paste0(cropped_loc, str_split_i(this.file,"[.]",1),"_cropped.csv"))
  
  #Create a dataframe of the raw and cropped data
  cropvraw <- left_join(this.data, cropped.data, by=c("row.num", "datetime")) %>%
    rename(raw.temp=temperature.x, cropped.temp =temperature.y) %>%#rename temperature from each file 
    #create new column of data type (raw or cropped for plotting in ggplot)
    pivot_longer(cols=raw.temp:cropped.temp, names_to="type", values_to="temp")
  
  cropvraw.plot <- ggplot(cropvraw, aes(x=datetime, y=temp, color=type)) +
    geom_line() +
    geom_point() +
    labs(title = paste0(" Raw versus Cropped data"),
         x = "Date", y = "Temperature (C)")+
    theme(axis.text = element_text(colour = "black", size = (12)))
  
  ggsave(paste0(croppedplots_loc, csv.site, "_rawvscroppeddata.png"), cropvraw.plot,
         width = 11, height = 8.5, units = "in")
  
}

cat("Done.", fill = TRUE)
```


## Demo of `crop_raw_data()`

```{r setup}
library(dataQCtools)
```


