---
title: "Making QC plots"
output:
  html_document:
    theme: flatly
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 3
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Editable script

Here is a script you can use if you would prefer to have all the code for the QC plot step in a single script rather than an installable package format. Just click the clipboard icon in the top right corner of this code chunk to copy the code, then open a new R script in RStudio and paste it there, and save the script wherever you would like to store it on your computer.

```{r, eval = FALSE}
# load necessary libraries
library(zoo)
library(grid)
library(plotly)
library(readxl)
library(scales)
library(runner) # for moving averages
library(ggthemes) # for plotting
library(tidyverse) # need for dplyr and ggplot2
library(lubridate) # for handling dates
library(gridExtra)
library(flextable) # for making tables with conditional formatting, merged cells and more
library(RColorBrewer)

# directory containing cropped data to be plotted
cropped_loc <- "data/2022_summer/2_cropped_csv/"
# directory where QC plots will be stored
qc_plots_loc <- "data/2022_summer/3_qc_plots/"

# get a list of all the cropped files
filenames <- list.files(path = cropped_loc, pattern=".csv")
# this is now a list of all filenames; we haven't read in the data yet, but
# make sure this lists all the cropped files you want to plot
filenames

# make QC plots!
make_qc_plots_of_cropped_data(cropped_loc, qc_plots_loc)
```


## Demo of `crop_raw_data()`

```{r setup}
library(dataQCtools)
```
